---
title: "LHI_BL_DISEASE.NODEAD_NO_DEAD"
author: "Tess Moriarty"
date: "07/08/2020"
output:
  html_document: default
  pdf_document: default
fig_width: 4
fig_text: 10
fig_height: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{R, include = FALSE}
# load packages

library(Hmisc)
library(plyr)
library(dplyr)
library(lme4)
library(ggplot2)
library(car)
library(MASS)
library(glmmTMB)
library(DHARMa)
library(multcomp)
library(emmeans)
library(RColorBrewer)
install.packages("tidyverse")
library(tidyverse)
library(vegan)
library(phyloseq)
library(viridis)
install.packages('unikn')
library('unikn')    
```

# load photo quadrat data

```{r}
getwd()
setwd("~/Documents/Tess_PhD/Lord Howe Island/Disease")
DISEASE.NODEAD <- read.csv("LHIDISEASER1nodead.csv", stringsAsFactors = FALSE, header = T, fill = TRUE)
DISEASE.NODEAD$Site <- as.factor(DISEASE.NODEAD$Site)
DISEASE.NODEAD$Date <- as.factor(DISEASE.NODEAD$Date)
DISEASE.NODEAD$Day <- as.factor(DISEASE.NODEAD$Day)
DISEASE.NODEAD$Month <- as.factor(DISEASE.NODEAD$Month)
DISEASE.NODEAD$Year <- as.factor(DISEASE.NODEAD$Year)
DISEASE.NODEAD$Growth.form <- as.factor(DISEASE.NODEAD$Growth.form)
DISEASE.NODEAD$Size <- as.factor(DISEASE.NODEAD$Size)
DISEASE.NODEAD$Transect <- as.factor(DISEASE.NODEAD$Transect)
DISEASE.NODEAD$Genera <- as.factor(DISEASE.NODEAD$Genera)
DISEASE.NODEAD$Taxa <- as.factor(DISEASE.NODEAD$Taxa)
DISEASE.NODEAD$Health.status <- as.factor(DISEASE.NODEAD$Health.status)
DISEASE.NODEAD$Unhealthy.status <- as.factor(DISEASE.NODEAD$Unhealthy.status)
DISEASE.NODEAD$Unhealthy <- as.factor(DISEASE.NODEAD$Unhealthy)
DISEASE.NODEAD$Unhealthy.ID <- as.factor(DISEASE.NODEAD$Unhealthy.ID)
str(DISEASE.NODEAD)
```


# subset data

```{r}
DISEASE.NODEAD.br <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Branching'),]
DISEASE.NODEAD.dig <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Digitate'),]
DISEASE.NODEAD.mas <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Massive'),]
DISEASE.NODEAD.sub <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Sub massive'),]
DISEASE.NODEAD.en <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Encrusting'),]
DISEASE.NODEAD.stag <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Staghorn'),]
DISEASE.NODEAD.tab <- DISEASE.NODEAD[which(DISEASE.NODEAD$Growth.form == 'Table'),]


DISEASE.NODEAD.BL <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Bleached'),]
DISEASE.NODEAD.COMP <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Competition'),]
DISEASE.NODEAD.HF <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Healthy'),]
DISEASE.NODEAD.OT <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Other'),]
DISEASE.NODEAD.DIS <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Discoloration'),]
DISEASE.NODEAD.DRUP <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Drupella'),]
DISEASE.NODEAD.FLO <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Fluoresence'),]
DISEASE.NODEAD.GA <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Growth anomaly'),]
DISEASE.NODEAD.INJ <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Injured'),]
DISEASE.NODEAD.PALE <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Pale'),]
DISEASE.NODEAD.PDEAD <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Partially Dead'),]
DISEASE.NODEAD.PRED <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Predation'),]
DISEASE.NODEAD.TISLOS <- DISEASE.NODEAD[which(DISEASE.NODEAD$Unhealthy.ID == 'Tissue loss'),]


DISEASE.NODEAD.stylo <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Stylophora'),]
#DISEASE.NODEAD.stylo
DISEASE.NODEAD.seri <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Seriatopora'),]
DISEASE.NODEAD.pdam <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Pocillopora'),]
DISEASE.NODEAD.Por <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Porites'),]
DISEASE.NODEAD.other <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Other'),]
DISEASE.NODEAD.iso <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Isopora'),]
DISEASE.NODEAD.acro <- DISEASE.NODEAD[which(DISEASE.NODEAD$Taxa == 'Acropora'),]

```

# try to make 'other' group of coral species

```{r}
# bar plots 

group.counts <- ddply(DISEASE.NODEAD, c("Site", "Month", "Transect"), summarise,  Acropora= sum(Taxa== "Acropora"), Isopora = sum(Taxa== "Isopora"), Other=sum(Taxa== "Other"), Pocillopora = sum(Taxa== "Pocillopora"), Porites = sum(Taxa== "Porites"), Seriatopora = sum(Taxa== "Seriatopora"), Stylophora = sum(Taxa== "Stylophora"))

#group.counts
group.counts.long <- gather(group.counts, Taxa, count,  Acropora:Stylophora)
group.counts.long$Taxa <- as.factor(group.counts.long$Taxa)
#group.counts.long
# use ddply to get %

group.percent <- ddply(DISEASE.NODEAD, c( "Site", "Month", "Transect"  ), summarise,  Acropora = sum(Taxa== "Acropora")/ length(Taxa) *100,  
                       Isopora = sum(Taxa== "Isopora")/ length(Taxa) *100, 
                       Other=sum(Taxa== "Other")/ length(Taxa) *100, 
                       Pocillopora = sum(Taxa== "Pocillopora")/ length(Taxa) *100, 
                       Porites = sum(Taxa== "Porites")/ length(Taxa) *100, 
                       Seriatopora = sum(Taxa== "Seriatopora")/ length(Taxa) *100, 
                       Stylophora = sum(Taxa== "Stylophora")/ length(Taxa) *100)

#group.percent
# Hstatus percent

group.percent.long <- gather(group.percent, Taxa, count, Acropora:Stylophora)

#group.counts.long
#group.percent.long
```


```{r}
#  broken down by Site + Month
group.percent.sum <- ddply(group.percent.long, c("Month" ,"Taxa"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N)
)
group.percent.sum

```

```{r}

stacked <- ggplot(group.percent.sum, aes(fill= Taxa, y=mean, x=Month)) + 
    geom_bar(position="fill", stat="identity")  + theme(axis.text.x = element_text(angle = 45, hjust = 0.8, size = 15), axis.text.y = element_text(angle = 0, hjust = 0.8, size = 15)) + labs(x = "Sample sites and months", y = "relative abundance (%)", title = "Taxa relative abundance between sites and months") +  theme(axis.title = element_text(size = 15)) + # axis title size
  theme(plot.title = element_text(size = 15)) # main title
stacked <- stacked + scale_x_discrete(name ="Month")

stacked

```
```{r}
#  broken down by Site + Month
group.percent.sum <- ddply(group.percent.long, c("Site","Month" ,"Taxa"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N)
)
group.percent.sum

```

__Relative abundance per site__
```{r}
# bar plots 

group.counts <- ddply(DISEASE.NODEAD, c("Site", "Transect"), summarise,  Acropora= sum(Taxa== "Acropora"), Isopora = sum(Taxa== "Isopora"), Other=sum(Taxa== "Other"), Pocillopora = sum(Taxa== "Pocillopora"), Porites = sum(Taxa== "Porites"), Seriatopora = sum(Taxa== "Seriatopora"), Stylophora = sum(Taxa== "Stylophora"))

#group.counts
group.counts.long <- gather(group.counts, Taxa, count,  Acropora:Stylophora)
group.counts.long$Taxa <- as.factor(group.counts.long$Taxa)
#group.counts.long
# use ddply to get %

group.percent <- ddply(DISEASE.NODEAD, c( "Site",  "Transect"  ), summarise,  Acropora = sum(Taxa== "Acropora")/ length(Taxa) *100,  
                       Isopora = sum(Taxa== "Isopora")/ length(Taxa) *100, 
                       Other=sum(Taxa== "Other")/ length(Taxa) *100, 
                       Pocillopora = sum(Taxa== "Pocillopora")/ length(Taxa) *100, 
                       Porites = sum(Taxa== "Porites")/ length(Taxa) *100, 
                       Seriatopora = sum(Taxa== "Seriatopora")/ length(Taxa) *100, 
                       Stylophora = sum(Taxa== "Stylophora")/ length(Taxa) *100)

#group.percent
# Hstatus percent

group.percent.long <- gather(group.percent, Taxa, count, Acropora:Stylophora)

#group.counts.long
#group.percent.long
```


```{r}
#  broken down by Site + Month
group.percent.sum <- ddply(group.percent.long, c("Site", "Taxa"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N)
)
group.percent.sum

```



```{r}

group.count.se <- ddply(group.percent.long, c("Site","Month" ,"Taxa"), summarise,
               N    = length(count),
               mean = mean(count),
               sd   = sd(count),
               se   = sd / sqrt(N)
)
group.count.se

abundance <- ggplot(group.count.se, aes(fill= Month, y=mean, x=Taxa)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) + theme_test() +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_fill_manual(values=c("#FFFF99", "#FF9933", "purple4")) + scale_y_continuous(limits = c(0, 60))  + ylab("Coral taxa relative abundance (%)")

abundance

```
# acropora
```{r}

acropor.relabun <- group.count.se
acropor.relabun <- dplyr::filter(group.count.se, Taxa != "Stylophora")

acropor.relabun <- group.count.se[which(group.count.se$Taxa == 'Acropora'),]
acropor.relabun <- ggplot(acropor.relabun , aes(fill= Month, y=mean, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Acropora relative abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

acropor.relabun

```
```{r}

acropor.relabun <- group.counts.long
acropor.relabun <- dplyr::filter(group.counts.long, Taxa != "Stylophora")

acropor.relabun <- group.counts.long[which(group.counts.long$Taxa == 'Acropora'),]
acropor.relabun <- ggplot(acropor.relabun , aes(fill= Month, y=count, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Acropora absolute abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

acropor.relabun

```

# isopora
```{r}

isopora.relabun <- group.count.se

isopora.relabun <- group.count.se[which(group.count.se$Taxa == 'Isopora'),]
iso.abundance <- ggplot(isopora.relabun , aes(fill= Month, y=mean, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Isopora relative abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

iso.abundance

```
```{r}

isopora.relabun <- group.counts.long

isopora.relabun <- group.counts.long[which(group.counts.long$Taxa == 'Isopora'),]
iso.abundance <- ggplot(isopora.relabun , aes(fill= Month, y=count, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Isopora absolute abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

iso.abundance

```

# pocillopora
```{r}

pdam.relabun <- group.count.se

pdam.relabun <- group.count.se[which(group.count.se$Taxa == 'Pocillopora'),]
pdam.abundance <- ggplot(pdam.relabun , aes(fill= Month, y=mean, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Pocillopora relative abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
pdam.abundance

```
# seriatopora
```{r}
seri.relabun <- group.count.se

seri.relabun <- group.count.se[which(group.count.se$Taxa == 'Seriatopora'),]
seri.abundance <- ggplot(seri.relabun , aes(fill= Month, y=mean, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Seriatopora relative abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
seri.abundance

```

# porites
```{r}

por.relabun <- group.count.se

por.relabun <- group.count.se[which(group.count.se$Taxa == 'Porites'),]
por.abundance <- ggplot(por.relabun , aes(fill= Month, y=mean, x=Month)) + 
    geom_bar(position="dodge", stat="identity")  + facet_wrap(~ Site) + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values=c("#C7F9EE", "#6DFDD2", "#1AC9E6")) + scale_y_continuous(limits = c(0, 60)) + theme_classic() + ylab("Porites relative abundance (%)") +theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
por.abundance

```




```{r}

stacked <- ggplot(group.percent.sum, aes(fill= Taxa, y=mean, x=Month)) + 
    geom_bar(position="fill", stat="identity")  + facet_wrap(~ Site) + theme(strip.text.x = element_text(angle = 45, hjust = 0.8, size = 18), axis.text.y = element_text(angle = 0, hjust = 0.8, size = 18)) + labs(x = "Sample sites and months", y = "Relative abundance (%)") +  theme(axis.title = element_text(size = 25)) + # axis title size
  theme(plot.title = element_text(size = 25)) # main title
stacked <- stacked + scale_x_discrete(name ="Month") + theme_classic(base_size = 15) + 
theme(plot.title = element_text(size = 15))  # main title 
 #ylab(expression("Dead" ~italic("Stylophora pistillata") ~"coral colonies (%)")) 
stacked <- stacked + scale_fill_discrete(name = "Coral taxa", labels = c(~ italic("Acropora spp."), ~italic ("Isopora cuneata"), "Other taxa", ~italic("Pocillopora damicornis"), ~italic("Porites spp."), ~ italic("Seriatopora hystrix"),~italic("Stylophora pistillata")))

stacked
```


```{r}

taxastacked <- ggplot(group.percent.sum, aes(fill=Taxa, y=mean, x=Month)) + facet_wrap(~Site) + 
    geom_bar(position="fill", stat="identity")+  theme(legend.title = element_blank())   + scale_fill_discrete(labels = c("Abiotic", "Algae", "Dead coral", "Other invertebrates", "Seagrass", "Soft coral", "Stony coral"))  +  scale_fill_discrete(name = "Coral taxa", labels = c(~ italic("Acropora spp."), ~italic ("Isopora cuneata"), "Other taxa", ~italic("Pocillopora damicornis"), ~italic("Porites spp."), ~ italic("Seriatopora hystrix"),~italic("Stylophora pistillata")))   + labs(x = "Site", y = "Coral taxa relative abundance (%)")  + # axis title size +
theme(panel.background = element_blank()) + # re label legend axis
scale_fill_viridis(discrete = TRUE, option = "B") # COSE OPTIONS BETWEEN a-d FOR DIFF COLOURS
barplot(1:7, col = magma(7))+
theme(plot.title = element_text(size = 10)) + # main title 
theme(axis.text.x = element_text(angle = 45, size = 10)) + theme(legend.text=element_text(size=10)) + theme(axis.text.y = element_text(size = 10), axis.title.x = element_text(size = 10, vjust = -0.8), axis.title.y = element_text(size = 10, vjust = 1)) 

taxastacked
```

```{r}

taxaabun <- ggplot(group.counts.long, aes(fill=Taxa, y=count, x=Month)) + facet_wrap(~Site) + 
    geom_bar(position="stack", stat="identity")+  theme(legend.title = element_blank())   + scale_fill_discrete(labels = c("Abiotic", "Algae", "Dead coral", "Other invertebrates", "Seagrass", "Soft coral", "Stony coral"))  +  scale_fill_discrete(name = "Coral taxa", labels = c(~ italic("Acropora spp."), ~italic ("Isopora cuneata"), "Other taxa", ~italic("Pocillopora damicornis"), ~italic("Porites spp."), ~ italic("Seriatopora hystrix"),~italic("Stylophora pistillata")))   + labs(x = "Site", y = "Coral taxa absolute abundance")  + # axis title size +
theme(panel.background = element_blank()) + # re label legend axis
scale_fill_viridis(discrete = TRUE, option = "B") # COSE OPTIONS BETWEEN a-d FOR DIFF COLOURS
barplot(1:7, col = magma(7))+
theme(plot.title = element_text(size = 10)) + # main title 
theme(axis.text.x = element_text(angle = 45, size = 10)) + theme(legend.text=element_text(size=10)) + theme(axis.text.y = element_text(size = 10), axis.title.x = element_text(size = 10, vjust = -0.8), axis.title.y = element_text(size = 10, vjust = 1)) 

taxaabun
```


# stats for taxa
```{r}

fit3 <- glmmTMB(Stylophora~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

plot(simulateResiduals(fit3))

```
# checking for pairwise comparisons   stylo bleached
```{r}
stylocomps <- emmeans(fit3, ~Site)

stylocomps1 <- emmeans(fit3, ~Month)

stylocomps4 <- emmeans(fit3, ~Month|Site)

pairs(stylocomps)

pairs(stylocomps1)

pairs(stylocomps4)
```
# stats for taxa
```{r}
# run a general linear model because none of my variables are independent 

#fit1 <- glmmTMB(Seriatopora~Site*Month  , family = poisson, data = group.counts)
#AIC(fit1)
#Anova(fit1)
#summary(fit1)

fit3 <- glmmTMB(Seriatopora~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Seriatopora)~ Site * Month, data = group.counts)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
#plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
# checking for pairwise comparisons   stylo bleached
```{r}
sericomps <- emmeans(sqrt.test3, ~Site)

sericomps1 <- emmeans(sqrt.test3, ~Month)

sericomps4 <- emmeans(sqrt.test3, ~Month|Site)

pairs(sericomps)

pairs(sericomps1)

pairs(sericomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Porites~Site*Month  , family = poisson, data = group.counts)

Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Porites~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

#sqrt.test3 <- aov(sqrt(Porites)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
porcomps <- emmeans(fit3, ~Site)

porcomps1 <- emmeans(fit3, ~Month)

porcomps4 <- emmeans(fit3, ~Month|Site)

pairs(porcomps)

pairs(porcomps1)

pairs(porcomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

#fit1 <- glmmTMB(Pocillopora~Site*Month  , family = poisson, data = group.counts)

#Anova(fit1)
#summary(fit1)

fit3 <- glmmTMB(Pocillopora~Site*Month+(1|Transect), family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

#sqrt.test3 <- aov(sqrt(Pocillopora)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
#plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
poccomps <- emmeans(fit3, ~Site)

poccomps1 <- emmeans(fit3, ~Month)

poccomps4 <- emmeans(fit3, ~Month|Site)

pairs(poccomps)

pairs(poccomps1)

pairs(poccomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Isopora~Site*Month +(1|Transect)  , family = poisson, data = group.counts)

Anova(fit1)
summary(fit1)

#fit3 <- glmmTMB(Isopora~Site*Month +(1|Transect), family = nbinom1, data = group.counts )
#Anova(fit3)
#summary(fit3)
#AIC(fit3)

#sqrt.test3 <- aov(sqrt(Isopora)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
#plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
isocomps <- emmeans(fit1, ~Site)

isocomps1 <- emmeans(fit1, ~Month)

isocomps4 <- emmeans(fit1, ~Month|Site)

pairs(isocomps)

pairs(isocomps1)

pairs(isocomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Acropora~Site*Month  , family = poisson, data = group.counts)
AIC(fit1)
Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Acropora~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Acropora)~ Site * Month, data = group.counts)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
plot(simulateResiduals(fit3))
plot(simulateResiduals(sqrt.test3))

# fit 1 - nbinom1 is the best fit 
```
```{r}
acrocomps <- emmeans(sqrt.test3, ~Site)

acrocomps1 <- emmeans(sqrt.test3, ~Month)

acrocomps4 <- emmeans(sqrt.test3, ~Month|Site)

pairs(acrocomps)

pairs(acrocomps1)

pairs(acrocomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Other~Site*Month  , family = poisson, data = group.counts)
AIC(fit1)
Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Other~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Other)~ Site * Month, data = group.counts)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
plot(simulateResiduals(fit3))
plot(simulateResiduals(sqrt.test3))
#plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```

```{r}
othercomps <- emmeans(fit3, ~Site)

othercomps1 <- emmeans(fit3, ~Month)

othercomps4 <- emmeans(fit3, ~Month|Site)

pairs(othercomps)

pairs(othercomps1)

pairs(othercomps4)
```


# Statistics between relative abundance of coral taxa 

```{r}

taxaabun <- ggplot(group.counts.long, aes(fill=Taxa, y=count, x=Month)) + facet_wrap(~Site) + 
    geom_bar(position="stack", stat="identity")+  theme(legend.title = element_blank())   + scale_fill_discrete(labels = c("Abiotic", "Algae", "Dead coral", "Other invertebrates", "Seagrass", "Soft coral", "Stony coral"))  +  scale_fill_discrete(name = "Coral taxa", labels = c(~ italic("Acropora spp."), ~italic ("Isopora cuneata"), "Other taxa", ~italic("Pocillopora damicornis"), ~italic("Porites spp."), ~ italic("Seriatopora hystrix"),~italic("Stylophora pistillata")))   + labs(x = "Site", y = "Coral taxa absolute abundance")  + # axis title size +
theme(panel.background = element_blank()) + # re label legend axis
scale_fill_viridis(discrete = TRUE, option = "B") # COSE OPTIONS BETWEEN a-d FOR DIFF COLOURS
barplot(1:7, col = magma(7))+
theme(plot.title = element_text(size = 10)) + # main title 
theme(axis.text.x = element_text(angle = 45, size = 10)) + theme(legend.text=element_text(size=10)) + theme(axis.text.y = element_text(size = 10), axis.title.x = element_text(size = 10, vjust = -0.8), axis.title.y = element_text(size = 10, vjust = 1)) 

taxaabun
```


# stats for taxa
```{r}

fit3 <- glmmTMB(Stylophora~Site*Month, family = nbinom1, data = group.percent )
Anova(fit3)
summary(fit3)
AIC(fit3)

plot(simulateResiduals(fit3))

```
# checking for pairwise comparisons   stylo bleached
```{r}
stylocomps <- emmeans(fit3, ~Site)

stylocomps1 <- emmeans(fit3, ~Month)

stylocomps4 <- emmeans(fit3, ~Month|Site)

pairs(stylocomps)

pairs(stylocomps1)

pairs(stylocomps4)
```
# stats for taxa
```{r}
# run a general linear model because none of my variables are independent 

#fit1 <- glmmTMB(Seriatopora~Site*Month  , family = poisson, data = group.percent)
#AIC(fit1)
#Anova(fit1)
#summary(fit1)

fit3 <- glmmTMB(Seriatopora~Site*Month, family = nbinom1, data = percent )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Seriatopora)~ Site * Month, data = percent)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
#plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
# checking for pairwise comparisons   stylo bleached
```{r}
sericomps <- emmeans(sqrt.test3, ~Site)

sericomps1 <- emmeans(sqrt.test3, ~Month)

sericomps4 <- emmeans(sqrt.test3, ~Month|Site)

pairs(sericomps)

pairs(sericomps1)

pairs(sericomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Porites~Site*Month  , family = poisson, data = group.counts)

Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Porites~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

#sqrt.test3 <- aov(sqrt(Porites)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
porcomps <- emmeans(fit3, ~Site)

porcomps1 <- emmeans(fit3, ~Month)

porcomps4 <- emmeans(fit3, ~Month|Site)

pairs(porcomps)

pairs(porcomps1)

pairs(porcomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

#fit1 <- glmmTMB(Pocillopora~Site*Month  , family = poisson, data = group.counts)

#Anova(fit1)
#summary(fit1)

fit3 <- glmmTMB(Pocillopora~Site*Month+(1|Transect), family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

#sqrt.test3 <- aov(sqrt(Pocillopora)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
#plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
poccomps <- emmeans(fit3, ~Site)

poccomps1 <- emmeans(fit3, ~Month)

poccomps4 <- emmeans(fit3, ~Month|Site)

pairs(poccomps)

pairs(poccomps1)

pairs(poccomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Isopora~Site*Month +(1|Transect)  , family = poisson, data = group.counts)

Anova(fit1)
summary(fit1)

#fit3 <- glmmTMB(Isopora~Site*Month +(1|Transect), family = nbinom1, data = group.counts )
#Anova(fit3)
#summary(fit3)
#AIC(fit3)

#sqrt.test3 <- aov(sqrt(Isopora)~ Site * Month, data = group.counts)
#plot(sqrt.test3)
#TukeyHSD(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
#plot(simulateResiduals(fit4))
#plot(simulateResiduals(sqrt.test3))
#plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```
```{r}
isocomps <- emmeans(fit1, ~Site)

isocomps1 <- emmeans(fit1, ~Month)

isocomps4 <- emmeans(fit1, ~Month|Site)

pairs(isocomps)

pairs(isocomps1)

pairs(isocomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Acropora~Site*Month  , family = poisson, data = group.counts)
AIC(fit1)
Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Acropora~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Acropora)~ Site * Month, data = group.counts)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
plot(simulateResiduals(fit3))
plot(simulateResiduals(sqrt.test3))

# fit 1 - nbinom1 is the best fit 
```
```{r}
acrocomps <- emmeans(sqrt.test3, ~Site)

acrocomps1 <- emmeans(sqrt.test3, ~Month)

acrocomps4 <- emmeans(sqrt.test3, ~Month|Site)

pairs(acrocomps)

pairs(acrocomps1)

pairs(acrocomps4)
```
```{r}
# run a general linear model because none of my variables are independent 

fit1 <- glmmTMB(Other~Site*Month  , family = poisson, data = group.counts)
AIC(fit1)
Anova(fit1)
summary(fit1)

fit3 <- glmmTMB(Other~Site*Month, family = nbinom1, data = group.counts )
Anova(fit3)
summary(fit3)
AIC(fit3)

sqrt.test3 <- aov(sqrt(Other)~ Site * Month, data = group.counts)
plot(sqrt.test3)
TukeyHSD(sqrt.test3)
AIC(sqrt.test3)
#fit2 <- glmmTMB(Dead~Site*Month + (1|Transect) , family = nbinom2, data = stylo.counts)
#Anova(fit2)

#fit3 <- glmmTMB(Dead~Site*Month + (1|Transect), family = poisson, data = stylo.counts)
#Anova(fit3)
#summary(fit3)

#fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = UNstatus.counts.bleached)
#Anova(fit4)
#summary(fit4)

#fit5 <- glmmTMB(Healthy~Month*Transect, family = poisson, data = UNstatus.counts.Healthy)
#Anova(fit5)
#summary(fit5)
plot(simulateResiduals(fit1))
plot(simulateResiduals(fit3))
plot(simulateResiduals(sqrt.test3))
#plot(simulateResiduals(fit3))
# fit 1 - nbinom1 is the best fit 
```

```{r}
othercomps <- emmeans(fit3, ~Site)

othercomps1 <- emmeans(fit3, ~Month)

othercomps4 <- emmeans(fit3, ~Month|Site)

pairs(othercomps)

pairs(othercomps1)

pairs(othercomps4)
```






```{r}
# run a general linear model because none of my variables are independent 

fit1<- glmmTMB(Bleached~Site*Month , family = nbinom1, data = B.percent)
fit1
Anova(fit1)
summary(fit1)
AIC(fit1)

fit2 <- glmmTMB(Bleached~Site*Month , family = nbinom2, data = B.percent)
Anova(fit2)
summary(fit2)
AIC(fit2)

fit3 <- glmmTMB(Bleached~Site*Month + (1|Transect), family = poisson, data = B.percent)
Anova(fit3)
summary(fit3)
AIC(fit3)

fit4 <- glmmTMB(Bleached~Site*Month, family = poisson, data = B.percent)
Anova(fit4)
summary(fit4)
AIC(fit4)

plot(simulateResiduals(fit1))
plot(simulateResiduals(fit2))
plot(simulateResiduals(fit3))
plot(simulateResiduals(fit4))

# fit 1 - is the best fit 
```


```{R}
b <- ggplot(group.percent.sum, aes(x =Taxa , y = mean, fill= Month))
b <- b + geom_bar(position = "dodge", stat = "identity", colour = "black")
b <- b + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) + facet_wrap(~ Site )
b <- b + theme_classic()
b <- b + theme(axis.text.x = element_text(angle = 90))
b <- b + scale_fill_manual("legend", values = c("T0" = "black", "T1" = "grey", "T2" = "white"))
b <- b + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title = "Hard coral Taxa abundance") + ylab("Coral colony (%)") + xlab("Coral taxa")

b
```

# Seriatopora abudnance
```{r}
# bar plots 

seri.counts <- ddply(DISEASE.NODEAD.seri, c("Site", "Month", "Transect", "Taxa"), summarise,  Healthy= sum(UNstatus== "Healthy"), Bleached = sum(UNstatus== "Bleached"), Other=sum(UNstatus== "Other"), Dead = sum(UNstatus== "Dead"))

seri.counts
seri.counts.long <- gather(seri.counts, UNstatus, count, Healthy, Bleached, Other, Dead)
seri.counts.long$UNstatus <- as.factor(seri.counts.long$UNstatus)
seri.counts.long
# use ddply to get %

seri.percent <- ddply(ECO.seri, c( "Site", "Month", "Transect" , "Taxa" ), summarise,  Healthy= sum(UNstatus== "Healthy")/ length(UNstatus) *100, Bleached = sum(UNstatus== "Bleached")/ length(UNstatus) *100, Other=sum(UNstatus== "Other")/ length(UNstatus) *100, Dead = sum(UNstatus== "Dead")/ length(UNstatus) *100)

seri.percent
# Hstatus percent

seri.percent.long <- gather(seri.percent, UNstatus, seri.counts, Healthy, Bleached, Other, Dead)

seri.counts.long

seri.percent.sum <- ddply(seri.percent.long, c("Site","Time" ,"UNstatus","Taxa"), summarise,
               N    = length(seri.counts),
               mean = mean(seri.counts),
               sd   = sd(seri.counts),
               se   = sd / sqrt(N)
)

seri.percent.sum
seri.percent.sum.bleached <- seri.percent.sum
seri.percent.sum.bleached$Site <- factor(seri.percent.sum.bleached$Site,levels = c("Sylphs Hole", "North Bay", "Coral Garden")) # Site location on graph
seri.percent.sum.bleached <- dplyr::filter(seri.percent.sum.bleached, UNstatus != "Other")
seri.percent.sum.bleached <- dplyr::filter(seri.percent.sum.bleached, UNstatus != "Healthy")
seri.percent.sum.bleached <- dplyr::filter(seri.percent.sum.bleached, UNstatus != "Dead")
b <- ggplot(seri.percent.sum.bleached, aes(x = UNstatus,y = mean, fill= Time))
b <- b + geom_bar(position = "dodge", stat = "identity", colour = "black")
b <- b + geom_errorbar(aes(ymin=mean-se, ymax = mean + se), width = 0.2, position = position_dodge(0.9)) + facet_wrap(~ Site )

b <- b + theme_classic()
b <- b + theme(axis.text.x = element_text(angle = 90))
b <- b + scale_fill_manual("legend", values = c("T0" = "white", "T1" = "white", "T2" = "white"))
b <- b + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title = "Seriatopora % bleached and dead coral colonies over time") + ylab("Bleached Seriatopora coral colonies (%)") + xlab("Time")

b

```